---
title: "Data Science"
author: "DSLovesR"
date: "December 22, 2019"
output: html_document
---
**CTRL+ALT+I**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Analysis

* There are six categories in which data analysis fall:
   + Descriptive
   + Exploratory
   + Inferential
   + Predictive
   + Causal
   + Mechanistic

## Descritive

The key goal of descritive anaylysis to **describe** or **summanries** a set of data. common descriptive analysis are central tendancy ( mean, median, mode) or measure of variability ( range, standard deviation , variance).


## Explarotary

Objective is to  examin or explore the data and find the **relationship** that were not known previousely. 


## Inferential

Objective is to use small sample of data to infer or same something about the **population** at large.

Inferential analysis is commonly the goal of statistical modelling, where you have a small amount of information to extrapolate and generalize that information to a larger group.

## Predictive

Objective is to use current data and make **prediction** about **future** data.
Essentially, you are using current and historical data to find patterns and predict the likelihood of future outcomes.


## Causal

Objective us to see the change in one variable when manipulate other variable- looking at the **cause** and **effect** of the **relationship**.


## Mechanistic

Objective of the mechanistic analysis is to understand the **exact changes in variables** that lead to **exact changes in other variables**.


-------------------------------------------------------------------
**Hyposthesis** is the **expected outcome** of an experiment.
For example- As height of child increase,  increase weight 


#Experimental Design

Experimental design is organizaing an experiment so that you have correct data to clearly and effectively answer your data science question.

**Confounder** An extraneous variable that may affect the relationship between dependent and independent variable.

**Control group**
**Placebo effect**- Subject knows the relationship which affect the result

**Blinding** to avoid placebo effect
**randomization**- balancing of confounders (**reduce confounders effect**)


**p-value**- Value that tells the probability the results of experiment were observed by chance.



#R Nuts and Bolts

R is dialect of **S**. S was developed by **John Chambers** in 1976 at Bell Labs for internal statistical analysis environment.


##Objects
 R has five basic or **atomic** classes of obejcts:
 
 * Character
 * Integer
 * Numeric ( real numbers)
 * complex
 * Logical 
 
The most basic object is **vector**.

##Attributes

R object can have following attributes:

* names, dimnames
* dimenions ( metric or array)
* class
* length
* Other user defined attributes/metadata
```{r}
x<-"Hello"
x
#change to numeric
x<- 100.11
print (x)

print(class(x))

```

Attributes of an object can be accessed using the **attributes()** function.

**c()** function can be used to create vectord of objects.
```{r}
x<- c( 1,2,3,4)
x
#sequence number usign :
x<- c(1:10)
print(x)
```

## Mixing Objects

When differnt objects are mixed in vector, **coercion** occures so that every element of vector is of the same class.  -*Implicit coercion*
```{r}
x <- c (1, TRUE, "Test")
#converts all elements to higher class objects
print(x)


```

*Explicit coercion* - one class to another class using as.* functions.
```{r}
x<- "Sample"
#explicit coercion- output NA in case converion is not possible.
print ( as.numeric(x))
x<- 100
print( as.character(x))
```

Elements of a list will have **double brackets [[]]** around them elements of other vectors just have the *single brackets[]*.
```{r}
v <- c(1:5)
print (v)

l <- list( 1, "Test", TRUE)
print (l)


```

**Factors** are used to represent categorical data.



```{r}
x<- factor( c("Male", "Female", "Female", "Male", "Neutral", "Male"))

print(x)

table(x)

unclass(x)

#order factor
y<- factor( c("Male", "Female", "Female", "Male", "Neutral", "Male"), levels = c("Male","Female","Neutral"))

print(y)

table(y)

unclass(y)

```


*Data.Frame(col1=, col2=)* and *names()*

#Reading Data
* Read.tables. read.csv for reading tabular data
* readlines
* load for reading in saved workspaces
* dget - reading in R code files
* source- reading in R code files
* unserialize - for reading single r objects in binary form

## dput-ting R object

```{r}
x <- data.frame( a1=1, b1="Test")
x
dput(x)

dput( x, file="dput01.R")

new.x <- dget("dput01.R")

new.x


```


#Connection interface

R has following connection to open the outside word
* file
* gzfile - compressed with  gzip 
* bzfile - compressed with bzip2 
* url - webpage

## Subsetting

* [ returns an object of the me class as the original
```{r}
x<- c("a", "b","c")
x[1]
x[x>"a"]

```

* [[ extract elements of a list or data frame.
```{r}
x<- list( foo = 1:4, bar =0.6)
x[1]
x[[1]]

x[[2]]

x$bar



```

* $ same as [[ but by **name**.

the **[[** operator can be used with computed indices; **$** can be used with literal names.

```{r}
x<- list( a=list(10, 11, 12), b= c( 11,22))
x[[c(1,3)]]
x[[1]][[3]]

x[[2]][[1]]
```


To preserve single element matrix use drop=FALSE

```{r}
x<- matrix(1:6,2,3)
#return vector 
x[1,2]

# to preserve matrix use DROP
x[1,2, drop=FALSE]

```

## partial matching

```{r}
x<- list ( partial = 1:5)
x$p

x[["p"]]

x[["p", exact=FALSE]]
```

## Remove NA values

First get logical vector by using is.na() and then use logical vector to return false value only

or complete.case() for multiple objects


# Binary operator 

       "Good" %p% "job!"

This will evaluate to: "Good job!"
```{r}
"%p%" <- function(args1,args2){ # Remember to add arguments!
   paste( args1, args2, sep=" ")
}
'I' %p% 'love' %p% 'R!'
```



Dates are represented by the 'Date' class and times are represented by the 'POSIXct' and 'POSIXlt'
| classes. Internally, dates are stored as the number of days since 1970-01-01 and times are stored as
| either the number of seconds since 1970-01-01 (for 'POSIXct') or a list of seconds, minutes, hours, etc.
| (for 'POSIXlt').

```{r}
t3<-"October 17, 1986 08:24"
t4<- strptime(t3, "%B %d, %Y %H:%M")
t4
 class(t4)

```


```{r}

t3<-"October 17, 2019 08:24"
t4<- strptime(t3, "%B %d, %Y %H:%M")

difftime( Sys.time(), t4, units='days')
```

